Efficient Machine Learning in a Memristor Network

### 2025-01-28 paper/paper.tex
Started implementing regression training.

\documentclass[reprint,superscriptaddress,prb,floatfix,showkeys]{revtex4-2}
\usepackage{empheq}
\usepackage{float}
\usepackage{lipsum}
\usepackage{mathtools, cuted}
\usepackage{esvect}
\usepackage{xcolor}
\usepackage[caption = false]{subfig}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{physics}
\usepackage[mathlines]{lineno}
\usepackage[capitalise]{cleveref}
\usepackage[english]{babel}
\usepackage{orcidlink}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amssymb}

@@

Artificial neural networks (ANNs) have become central to many technological advancements, achieving remarkable success in tasks ranging from image recognition to natural language processing. These networks typically operate in two phases: inference and backpropagation. The energy requirements of ANNs has grown significantly, negatively impacting the efficiency and scalability of ANNs \cite{energy_cost}.

@@

While the inference energy consumption has been addressed by hardware accelerators, the backpropagation still remains a problem. As an alternative to ANN performed strictly in-silico, physical neural networks (PNN) have been introduced, where physical processes rather than mathematical operations are trained \cite{}. In fact, physical processes are often faster and consume less energy than their digital counterpart.

@@

The PNN we consider is an electrical network make of nodes connected by resistive elements. As resistive elements, we consider conical microfluidic channels that are schematically represented in \cref{fig:memristor_scheme}. 
\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{plots/memristor/memristor_scheme.pdf}
    \caption{A schematic representation of the conical fluidic memristor. A cone of lenght $L$, base radius $R_b$ and tip radius $R_t$ with charged surface connnects two bulk reservoirs of a 1:1 electrolyte in an acqueous solution. The bulk density in the base $\rho_b$ differs from the bulk density in the tip $\rho_t$. At the far side of the memristor, an electrical potential and pressure drop are applied. Overall, this results in a chemical, electrical and mechanical sensistive system.}
    \label{fig:memristor_scheme}
\end{figure} 
An azimuthally symmetric conical channel connects two reservoirs that contain an incompressible aqueous 1:1 electrolyte with viscosity $\eta=1.01mPa s$ containing ions with diffusion coefficients $D = 1.75\mu m^2 ms^{-1}$ and charge $\pm e$ with $e$ the proton charge. The cone has a base radius of $R_b = 200$nm, a tip radius of $R_t=50$nm and has a length $L = 10\mu$m, which sets it in the long channel limit where entrance and exit effects can be neglected. The walls of the channel are homogeneously charged and generate a surface potential $\psi_0$ that attracts the opposite-sign ions present in the solution due to the dissociation of the electrolyte. The electric potential decreases exponentially away from the surface due to the electrostatic screening of the attracted ions: an electric double layer (EDL) forms on the charged surfaces.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\columnwidth]{plots/voltage_divider/vd_scheme.pdf}
    \caption{}
    \label{fig:vd_scheme}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{plots/voltage_divider/mse_weights_vd.pdf}
    \caption{}
    \label{fig:mse_weights_vd}
\end{figure}

>>>>

\documentclass[reprint,superscriptaddress,prb,showkeys]{revtex4-2}
\usepackage{empheq}
\usepackage{float}
\usepackage{lipsum}
\usepackage{mathtools, cuted}
\usepackage{esvect}
\usepackage{xcolor}
\usepackage[caption = false]{subfig}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{physics}
\usepackage[mathlines]{lineno}
\usepackage[capitalise]{cleveref}
\usepackage[english]{babel}
\usepackage{orcidlink}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amssymb}

@@

Artificial neural networks (ANNs) have become central to many technological advancements, achieving remarkable success in tasks ranging from image recognition to natural language processing. These networks typically operate in two phases: inference and backpropagation. The energy requirements of ANNs has grown significantly, negatively impacting the efficiency and scalability of ANNs\cite{energy_cost}.

@@

While the inference energy consumption has been addressed by hardware accelerators, the backpropagation still remains a problem. As an alternative to ANN performed strictly in-silico, physical neural networks (PNN) have been introduced, where physical processes rather than mathematical operations are trained\cite{}. In fact, physical processes are often faster and consume less energy than their digital counterpart.

@@

The PNN we consider is an electrical network make of nodes connected by resistive elements. As resistive elements, we consider conical microfluidic channels that are schematically represented in \cref{fig:memristor_scheme}. 
\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{plots/memristor/memristor_scheme.pdf}
    \caption{A schematic representation of the conical fluidic memristor. A cone of lenght $L$, base radius $R_b$ and tip radius $R_t$ with charged surface connnects two bulk reservoirs of a 1:1 electrolyte in an acqueous solution. The bulk density in the base $\rho_b$ differs from the bulk density in the tip $\rho_t$. At the far side of the memristor, an electrical potential and pressure drop are applied. Overall, this results in a chemical, electrical and mechanical sensistive system.}\label{fig:memristor_scheme}
\end{figure} 
An azimuthally symmetric conical channel connects two reservoirs that contain an incompressible aqueous 1:1 electrolyte with viscosity $\eta=1.01mPa s$ containing ions with diffusion coefficients $D = 1.75\mu m^2 ms^{-1}$ and charge $\pm e$ with $e$ the proton charge. The cone has a base radius of $R_b = 200$nm, a tip radius of $R_t=50$nm and has a length $L = 10\mu$m, which sets it in the long channel limit where entrance and exit effects can be neglected. The walls of the channel are homogeneously charged and generate a surface potential $\psi_0$ that attracts the opposite-sign ions present in the solution due to the dissociation of the electrolyte. The electric potential decreases exponentially away from the surface due to the electrostatic screening of the attracted ions: an electric double layer (EDL) forms on the charged surfaces.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\columnwidth]{plots/voltage_divider/vd_scheme.pdf}
    \caption{}\label{fig:vd_scheme}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{plots/voltage_divider/mse_weights_vd.pdf}
    \caption{}\label{fig:mse_weights_vd}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\columnwidth]{plots/voltage_divider/evolution_targets.pdf}
    \caption{}\label{fig:evolution_targets}
\end{figure}

\cleardoublepage

\subsection{Training a more general network}

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{plots/general_network/mse_general.pdf}
    \caption{Network to the right has inputs $[5,0]$ nad desired outputs $[3, 4]$, delta weights a third of their intial value and learning rates $1e^{-5}, 1e^{-4}, 1e^{2}$. The network to the right has $[5,1,0]$ and output $[3,4,2]$}\label{fig:mse_general}
\end{figure}
